---
title: "Maps"
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r}
#| message: false
#| warning: false

library(sf)
states <- read_sf("https://rstudio.github.io/leaflet/json/us-states.geojson") 

library(leaflet)
library(dplyr)
library(tidyverse)
```


2.  The `states` dataset in the `poliscidata` package contains 135 variables on each of the 50 US states. See [here](https://rdrr.io/cran/poliscidata/man/states.html) for more detail.

Your task is to create a *two* meaningful choropleth plots, one using a numeric variable and one using a categorical variable from `poliscidata::states`. You should make *two* versions of each plot: a static plot using the `maps` package and `ggplot()`, and an interactive plot using the `sf` package and `leaflet()`. Write a sentence or two describing what you can learn from each plot.

Here's some R code and hints to get you going:

```{r}
#| echo: false

polisci_hw4 <- poliscidata::states |>
  as_tibble() |>
  mutate(states = str_to_lower(str_squish(as.character(state)))) |>
  select(states, college, cook_index3)

states_poly_hw4 <- as_tibble(map_data("state")) |>
  mutate(states = str_to_lower(region)) |>
  left_join(polisci_hw4, by = "states")

states_sf_clean <- read_sf("https://rstudio.github.io/leaflet/json/us-states.geojson") |>
  mutate(state_key = str_to_lower(name),
         states = state_key) |>
  filter(!name %in% c("Alaska", "Hawaii", "Puerto Rico")) |>
  left_join(polisci_hw4, by = "states")
```

static numeric variable :

```{r}
#| echo: false 

ggplot(states_poly_hw4,
       aes(x = long, y = lat, group = group, fill = college)) +
  geom_polygon(color = "white", linewidth = 0.2) +
  coord_map("polyconic") +
  scale_fill_viridis_c( name = "college index") +
  labs(title = "percentage of people who attend college")
```

from this map we can learn that majority of states have between 20% and 35% college attendance

interactive numeric variable :

```{r}
#| echo: false

num_pal <- colorNumeric("YlOrRd", domain = states_sf_clean$college)

leaflet(states_sf_clean) |>
  addPolygons(
    fillColor = ~num_pal(college),
    weight = 1, color = "white", 
    fillOpacity = 0.7,
    label = ~paste0(name, college),
   highlightOptions = highlightOptions(   
      weight = 3, 
      color = "black", 
      bringToFront = TRUE
    )
  )
```

This map is very helpful because as a non-American who is not familiar with state names, I get to know the state name and see the percentage college attendance as I interact with the map

static categorical varivale

```{r}
#| echo: false

ggplot(states_poly_hw4, aes(x = long, y = lat, group = group, fill = cook_index3)) +
  geom_polygon(color = "white", linewidth = 0.2) +
  coord_map("polyconic") +
  scale_fill_manual(values = c("Dem" = "blue", "Even" = "purple", "Rep" = "red"), 
                    name = "Partisanship") +
  theme_void() +
  labs(title = "US States Cook partisan index")
```

from this map, we can see that majority of the US states do not have a prefered cook party

intercative categorical variable

```{r}
#| echo: false

cat_pal <- colorFactor( palette = c("blue", "purple", "red"), 
  domain = states_sf_clean$cook_index3
)

leaflet(states_sf_clean) |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addPolygons(
    fillColor = ~cat_pal(cook_index3),
    weight = 1, color = "white", fillOpacity = 0.7,
    label = ~paste0(name, ": ", cook_index3),
    highlightOptions = highlightOptions(weight = 2, color = "#666", bringToFront = TRUE)
  ) |>
  addLegend(pal = cat_pal, values = ~cook_index3, title = "Political Lean")
```

from this graph we get to interact with the plot and see where the people actually lean towards in term of political party
