---
title: "Maps"
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: false 
#| message: false
#| warning: false
#| echo: false

library(sf)
states <- read_sf("https://rstudio.github.io/leaflet/json/us-states.geojson") 

library(leaflet)
library(dplyr)
library(tidyverse)
```


```{r}
#| echo: false

polisci_hw4 <- poliscidata::states |>
  as_tibble() |>
  mutate(states = str_to_lower(str_squish(as.character(state)))) |>
  select(states, college, pot_policy)

states_poly_hw4 <- as_tibble(map_data("state")) |>
  mutate(states = str_to_lower(region)) |>
  left_join(polisci_hw4, by = "states")

states_sf_clean <- read_sf("https://rstudio.github.io/leaflet/json/us-states.geojson") |>
  mutate(state_key = str_to_lower(name),
         states = state_key) |>
  filter(!name %in% c("Alaska", "Hawaii", "Puerto Rico")) |>
  left_join(polisci_hw4, by = "states")
```

## Numeric variable :college attendance index

```{r}
#| echo: false 

ggplot(states_poly_hw4,
       aes(x = long, y = lat, group = group, fill = college)) +
  geom_polygon(color = "white", linewidth = 0.2) +
  coord_map("polyconic") +
  scale_fill_viridis_c( name = "college index") +
  labs(title = "percentage of people who attend college per state")
```
- from this map we can learn that majority of states have between 20% and 35% college attendance



```{r}
#| echo: false

num_pal <- colorNumeric("YlOrRd", domain = states_sf_clean$college)

leaflet(states_sf_clean) |>
  setView(-96, 37.8, 3) |>
  addTiles() |>
  addPolygons(
    fillColor = ~num_pal(college),
    weight = 1, color = "white", 
    fillOpacity = 0.7,
    label = ~paste0(name," : ", college,"%"),
   highlightOptions = highlightOptions(   
      weight = 3, 
      color = "black", 
      bringToFront = TRUE
    )
  )
```

- This map is very helpful because as a non-American who is not familiar with state names, I get to know the state name and see the percentage college attendance as I interact with the map

## Categorical variable : laws on marijuana consumption

```{r}
#| echo: false

ggplot(states_poly_hw4, aes(x = long, y = lat, group = group, fill = pot_policy)) +
  geom_polygon(color = "white", linewidth = 0.2) +
  coord_map("polyconic") +
  theme_void() +
  labs(title = "US States laws on marijuana")
```
- From this graph we get to see each states legal laws on marijuana consumption. we can tell from the map that the marijuana is decriminalized in most regions except for the south where it is not legal at all.

```{r}
#| echo: false

cat_pal <- colorFactor( palette = c("blue", "purple", "red"), 
  domain = states_sf_clean$pot_policy
)

leaflet(states_sf_clean) |>
   setView(-96, 37.8, 3) |>
  addTiles() |>
  addPolygons(
    fillColor = ~cat_pal(pot_policy),
    weight = 1, color = "white", fillOpacity = 0.7,
    label = ~paste0(name, ": ", pot_policy),
    highlightOptions = highlightOptions(weight = 3, color = "black", 
                                        bringToFront = TRUE)
  ) |>
  addLegend(pal = cat_pal, values = ~pot_policy, title = "view on marijuana")
```

- From this graph we get to interact with the plot and see each states legal laws on marijuana consumption. we can tell from the map that the marijuana is decriminalized in most regions except for the south where it is not legal at all.
